Anxiety, avoidance, and sequential evaluation

### 2019-12-05 manuscripts/manuscript/manuscript.tex
Updates from Overleaf

Though anxiety disorders differ in their particular symptomology, and in the content and situations that elicit symptoms, they all are similarly characterized by aberrations in the processing of and response to threat \citep{dsm5}. In particular, at least three symptoms manifest across many of the anxiety disorders. First, anxiety is associated with exaggerated threat appraisal, or a bias towards evaluating threat as disproportionately greater in likelihood and severity than is warranted \citep{ClarkBeck2011}. Second, anxiety is also associated with fear generalization, wherein the primary threat becomes associated with increasingly distal locations, events, and thoughts \citep{dymond2015}. Finally, anxiety is associated with persistent avoidance behavior, which often occurs well in advance of the materialization of actual threat \citep{Arnaudova2017}. Excessive avoidance behaviors are an especially harmful aspect of anxiety disorders, both because they interfere with daily life and because they indirectly maintain anxiety by preventing learning from the non-occurrence of perceived threats. Though laboratory studies of decision making and learning in anxious populations have corroborated these clinical observations \citep{Harle2017, norbury2018, Aylward2019}, none offer an explanation as to the root of these symptoms.

@@

For any particular state-action $(s,a)$, this is necessarily defined only relative to a policy $\pi(a' \mid s')$ specifying the assumed distribution of \emph{future} choices. The return can be optimized self-consistently under the assumption that the agent makes the return-maximizing choice at each step in the future, leading to the familiar expression for the optimal values,

@@

The ``max'' operator in Eq.~\ref{eq:bellman} yields a fundamental asymmetry between approach and avoidance, as illustrated in Fig.~1. It formalizes the assumption that the agent makes the return-maximizing choice at each step. Through this operation, opportunities for reward propagate (which maximize the argument) recursively to earlier steps, but avoidable dangers do not, because the return-maximizing action is to avoid. (To the extent obtaining reward or avoiding harm are only imperfectly achievable, this propagation and attenuation at each step are graded according to success probability, but the basic asymmetry remains.) This principle is highlighted in a toy MDP (Fig.~1): a deterministic open gridworld with two terminal states, a rewarding state and a punishing one (Fig.~1a). The optimal state values $V^* = \max_a Q^*(s,a)$ (Fig.~1b) reflect a ``mountain'' of opportunity propagating recursively from the reward. Conversely, because harm is avoidable in this environment, its negative value is contained: all states (even those adjacent to threat) represent the positive opportunity for reward.

@@

The parameter $w$ controls the degree of pessimism. An optimistic agent ($w = 1$) expects in the future to act fully in accordance with its preferences, whereas a pessimistic agent ($w = 0$) expects to act contrary to its preferences. Importantly, this belief is false under the model: that is, we assume, at each step, that the agent actually chooses by maximizing the values given by Equation~\ref{eq:pessimism}. But these values are always computed under the assumption that the agent will then not maximize at all later steps. (Another way to say this is that we assume choices at each step are optimal, but only under the assumption that later choices will not be.) Fig.~1c,d illustrates the the consequences of different levels of pessimism for valuation in an example gridworld. The value of threat propagates, with increasing distance across the state space, due to an increasing expectation that the agent may fail to avoid threat in the future.

@@

In what follows, we will demonstrate through simulation how our simple model can account for anxious behavior in laboratory-based studies of sequential learning and decision making. We will also show that our model is consistent with clinical theory describing the transition from clinical anxiety to depression. Unless otherwise noted, state and action values under varying degrees of pessimism were solved for using the value iteration dynamic programming method\citep{SuttonBarto2018}. All simulations were implemented in the Python programming language and the code is publicly available at \url{https://github.com/ndawlab/seqanx}.

@@

\begin{figure}
  \centerline{%
    \resizebox{1.0\textwidth}{!}{\includegraphics[trim={0 0 0 0},clip]{../../figures/02_appavo.png}}
  }
  \par \textbf{Figure 2:} (A) The Balloon Analog Risk Task (BART)\citep{Lejuez2002,Maner2007, ramirez2015}. The risk of balloon burst (point loss) increases with each pump. The optimal policy (number of pumps) under increasingly pessimistic valuation is presented for low (B) and high risk (C) balloons. The optimistic agent ($w=1$) prefers a policy reflecting the true environmental risk. The moderate ($w=0.6$) and strongly ($w=0.2$) pessimistic agents cash-out earlier, as is observed in anxious individuals. (D,E) In the sleeping predator task, the risk of loss is constant but the cost of loss still increases as more rewards are gathered. The value of reward pursuit under increasingly pessimistic valuation is presented for scenarios with low (D) and high risk (E)  of predator awakening. The relative value of approach (vs. avoid) decreases with loss amount and threat level, and moreso under pessimistic assumptions. (Parameters: $\gamma = 1.0$)
\end{figure}

@@

As shown in Fig.~2, our model easily accommodates this result. Whereas optimistic agents pump until the marginal gain of a pump no longer offsets the chance of the balloon bursting, optimal choice under increasingly pessimistic (i.e., anxious) assumptions cashes out progressively earlier -- similar to empirical findings\citep{Maner2007, ramirez2015}. This is because it anticipates and avoids future errors in choice, which would otherwise result in the balloon popping. Our model can analogously explain other manifestations of biased approach-avoid conflict in anxiety\cite{fung2019}.

@@

Relatedly, the model can also capture findings of increased behavioral inhibition, measured as prolonged response times, in anxious individuals under threat\cite{bach2015}. In the behavioral inhibition task, participants seek tokens adjacent to a virtual "sleeping predator", which are all lost if the predator awakes. Though in this variant the risk of predation is constant throughout a trial (rather than increasing, as in the BART) the potential loss from capture still increases with each token collected. Bach \cite{bach2015} finds that participants are slower to collect tokens as this potential loss increases, and that this slowing is enhanced by subclinical anxiety. We can capture this effect in our model by noting that the relative value of approach compared to avoidance is reduced as potential loss and the risk that the predator awakes increase (Fig.~2d/e). These effects are amplified under more pessimistic (i.e., in our model, more anxious) assumptions about future actions. Thus, as before, anxiety in our model produces greater and earlier choice of avoidance, analogous to earlier cash-out in the BART. To extend this effect to reaction times, we adopt the further, standard assumption that actions (here, approach) are slower when their relative value compared to alternatives is lower \citep{oud2016}. In this case, the model captures behavioral inhibition (slower responses as threat increases), and its enhancement by anxiety, as measured by Bach. (Note that we need not assume the coupling of reaction times to action value spread is due to difficulty in decision formation per se, which Bach argues against: it may, for instance, reflect Pavlovian initiation biases \cite{niv2007}.)

@@

Another laboratory phenomenon associated with anxiety is ``aversive pruning'' in planning \cite{Huys2012, Lally2017}. This refers to the idea that when evaluating future action trajectories in a sequential task like chess, people are resource-limited, cannot evaluate all possible options, and must selectively consider certain paths and neglect others. One proposal for how people accomplish this is aversive pruning \citep{Huys2012}, wherein choice sequences involving large losses are discarded from further evaluation. An example of aversive pruning is shown in Fig.~3a. Although the optimal choice in the decision tree from is to weather an initial large loss (e.g., -70) in order to reap the large gain that follows, people tend to disfavor this path suggesting they prune it and consequently neglect the later gain. The degree of such pruning correlates with subclinical depressive \cite{Huys2012} and anxiety \cite{Lally2017} symptoms.

Our model predicts this result (Fig.~3b), though for a somewhat different reason. In our model, pessimistic (anxious) agents neglect large gains deeper in the tree, not because they fail to consider them (here we assume full evaluation of the Bellman equation), but because with increasing anxiety they increasingly expect the potential of choosing incorrectly afterward, thus failing to recoup the loss (and, mathematically, probabilistically pruning the better branches). Future research could use slight variants in the decision trees to tease apart these different interpretations. Tree order matters in delineating between these theories. What's one good example?

@@

Of note in this respect, anxiety and depression are highly comorbid, with almost half of individuals with a lifetime depression diagnosis also diagnosed with an anxiety disorder \citep{kessler2015}. One notable proposal is that this association often arises longitudinally: in particular, that clinical anxiety precedes certain types of depression \citep{alloy1990, jacobson2014}. The idea, in brief, is that uncertainty in one's ability in the face of future threat results in anxiety and avoidance behaviors. Persistent avoidance, in turn, begets foregone reward, and ultimately to a belief that reward is unobtainable and subsequently depression. This story is borne out by simulations of learning in our model (Fig.~4) in environments like that of Fig.~1. Over the course of learning, the penumbra of negative value under pessimistic assumptions spreads gradually throughout the environment. This can in turn lead the agent to expect no reward and, also echoing the anergic symptoms of depression, forego action altogether.

In addition to accounting for the comorbidity of anxiety and depression, our model also hints at a reason for the longevity and recurrence of anxiety disorders even with treatment. Because pessimistic expectations allow for threat value to spread to states and actions far antecedent of the primary danger (e.g., Fig.~1d), it would accordingly also take a great many steps of iterative learning to correct all these exaggerated appraisals of threat. Frustratingly, these biased estimates of value may still remain even after a misbelief in the efficacy of future action is corrected for in a course of therapy. This phenomenon (similar to failures of model-free RL algorithms to adjust to reward revaluation without extensive relearning; \cite{daw2005}) may offer at least a partial answer to a classic puzzle in pathological avoidance, i.e. why it is so resistant to extinction \cite{moutoussis2018}, and to the unfortunately high rates of anxiety recurrence following treatment\citep{pittig2018}.

@@

We are not the first to propose a formal theory of control in psychiatry using MDPs. Huys \& Dayan \cite{HuysDayan2009} also provided a computational account of learned helplessness through simple models of one-step action-outcome contingencies. Our accounts differ particularly in our exclusive focus on control in the sequential setting, which Huys \& Dayan did not address. Indeed, we propose that ultimately key to anxiety is precisely the way in which evaluation in sequential tasks is necessarily reliant on expectations about future choice and events. Similarly, research and modeling by Bishop and colleagues \citep{browning2015, gagne2018} also taking a decision theoretic approach has stressed the importance of uncertainty, and aversion to it, as a core feature of anxiety. As for control, we extend this view to focus on how uncertainty is resolved in the sequential setting, and also to zero in on particular instances of uncertainty (about future actions, and some other options discussed next) and misbeliefs about them that, we argue, are particularly consequential.

@@

Insert paragraph here about testing. Delgado experiment for one. Learning transitions another?

@@

@@

It is important to note that the present model may not describe all anxiety disorders with equal accuracy. Indeed, our analysis of pessimistic sequential evaluation is, by definition, a model of prospective cognition. Thus, the present results are more likely to accurately describe the anxiety disorders which primarily involve aberrations in future-oriented cognitive processes, such as generalized and social anxiety disorders. Naturally, the present model can account neither for the compulsive behaviors of obsessive-compulsive disorder (OCD) nor the memory disturbances of post-traumatic stress disorder (PTSD). That said, recent clinical studies suggest that diminished perceived control is a vulnerability factor common to all anxiety disorders, including OCD and PTSD \citep{gallagher2014a, gallagher2014b}. Importantly, aspects of other psychiatric disorders that involve future-oriented misbeliefs, worry, and avoidance behaviors (e.g. eating disorders\citep{konstantellou2011}) may similarly be well-described by the current account. Speculatively, the model may also have implications for understanding dysregulated goal pursuit in mania and bipolar disorder. Manic symptoms are associated with increases in perceived control \citep{
implications for understanding increased perceived control in mania \cite{}. As such, the present results are transdiagnostic and not limited to one particular diagnosis. Finally, the relationship between evaluation, planning, and neural replay discussed above suggests potential future work that might help to bring this model into contact with the more memory-related aspects of PTSD. For instance, the same replay processes that can be used to evaluate actions can also, in theory, update predictive representations, cognitive maps, or models of the environment \citep{russek_predictive_2017}, such as the successor representation \cite{dayan1993, momennejad_successor_2017}. If so, similar biases in replay could result in not just aberrant avoidance behavior, but also progressive, aberrant remodeling of world models or cognitive maps, an observation which may connect to the rich and complex set of issues on memory involvement in PTSD.

@@

\textbf{Averse Pruning} In the aversive pruning task \citep{Huys2012, Lally2017}, participants learn to navigate a six-state graphworld, where each state is directly connected to only two other states. Each state is associated with some reward (cost), and participants must plan trajectories through the state-space so as to maximize reward (minimize cost).

>>>>

Though anxiety disorders differ in their particular symptomology, and in the content and situations that elicit symptoms, they all are similarly characterized by aberrations in the processing of and response to threat \citep{dsm5}. In particular, at least three symptoms manifest across many of the anxiety disorders. First, anxiety is associated with exaggerated threat appraisal, or a bias towards evaluating threat as disproportionately greater in likelihood and severity than is warranted \citep{ClarkBeck2011}. Second, anxiety is also associated with fear generalization, wherein the primary threat becomes associated with increasingly distal locations, events, and thoughts \citep{dymond2015}. Finally, anxiety is associated with persistent avoidance behavior, which often occurs well in advance of the materialization of actual threat \citep{Arnaudova2017}. (Here we distinguish between avoidance and escape, where the former describes actions taken to prevent the onset of threats whereas the latter describes defensive responses to proximal threat.) Excessive avoidance behaviors are an especially harmful aspect of anxiety disorders, both because they interfere with daily life and because they indirectly maintain anxiety by preventing learning from the non-occurrence of perceived threats. Though laboratory studies of decision making and learning in anxious populations have corroborated these clinical observations \citep{Harle2017, norbury2018, Aylward2019}, none offer an explanation as to the root of these symptoms.

@@

For any particular state-action $(s,a)$, this is necessarily defined relative to a policy $\pi(a' \mid s')$ specifying the assumed distribution of \emph{future} choices. (It is also defined relative to a discount rate $\gamma$, controlling the present value of future outcomes.) The return can be optimized self-consistently under the assumption that the agent makes the return-maximizing choice at each step in the future, leading to the familiar expression for the optimal values,

@@

The ``max'' operator in Eq.~\ref{eq:bellman} yields a fundamental asymmetry between approach and avoidance, as illustrated in Fig.~1. It formalizes the assumption that the agent makes the return-maximizing choice at each step. Through this operation, opportunities for reward (which maximize the argument) propagate recursively to earlier steps, but avoidable dangers do not, because the return-maximizing action is to avoid. (To the extent obtaining reward or avoiding harm are only imperfectly achievable, this propagation and attenuation at each step are graded according to success probability, but the basic asymmetry remains.) This principle is highlighted in a toy MDP (Fig.~1): a deterministic open gridworld with two terminal states, a rewarding state and a punishing one (Fig.~1a). The optimal state values $V^* = \max_a Q^*(s,a)$ (Fig.~1b) reflect a ``mountain'' of opportunity propagating recursively from the reward. Conversely, because harm is avoidable in this environment, its negative value is contained: all states (even those adjacent to threat) represent the positive opportunity for reward.

@@

The weight parameter $w$ controls the degree of pessimism. An optimistic agent ($w = 1$) expects in the future to act fully in accordance with its preferences, whereas a pessimistic agent ($w = 0$) expects to act contrary to its preferences. Importantly, this belief is false under the model: that is, we assume, at each step, that the agent actually chooses by maximizing the values given by Equation~\ref{eq:pessimism}. But these values are always computed under the assumption that the agent will then not maximize at all later steps. (Another way to say this is that we assume choices at each step are optimal, but only under the assumption that later choices will not be.) Fig.~1c,d illustrates the the consequences of different levels of pessimism for valuation in an example gridworld. The value of threat propagates, with increasing distance across the state space, due to an increasing expectation that the agent may fail to avoid threat in the future.

@@

In what follows, we will demonstrate through simulation how our simple model can account for anxious behavior in laboratory-based studies of \emph{sequential learning}  and decision making. (Because in our model anxiety arises through biased sequential evaluation, we will not address one-step bandit tasks where others have reported learning deficits associated with anxiety \citep{Harle2017, Aylward2019}.) We will also show that our model is consistent with clinical theory describing the transition from clinical anxiety to depression. Unless otherwise noted, state and action values under varying degrees of pessimism were solved for using the value iteration dynamic programming method\citep{SuttonBarto2018}. All simulations were implemented in the Python programming language and the code is publicly available at \url{https://github.com/ndawlab/seqanx}.

@@

\begin{figure}
  \centerline{%
    \resizebox{1.0\textwidth}{!}{\includegraphics[trim={0 0 0 0},clip]{../../figures/02_appavo.png}}
  }
  \par \textbf{Figure 2:} (A) The Balloon Analog Risk Task (BART)\citep{Lejuez2002,Maner2007, ramirez2015}. The risk of balloon burst (point loss) increases with each pump and does so earlier for the high risk (red) than low risk (blue) balloons. (For full rules of the task, see Methods.) The optimal policy (number of pumps) under increasingly pessimistic valuation is presented for low (B) and high risk (C) balloons. The optimistic agent ($w=1$) prefers a policy reflecting the true environmental risk. The moderate ($w=0.6$) and strongly ($w=0.2$) pessimistic agents cash-out earlier, as is observed in anxious individuals. (D,E) In the sleeping predator task, the risk of loss is constant but the cost of loss still increases as more rewards are gathered. The value of reward pursuit under increasingly pessimistic valuation is presented for scenarios with low (D) and high risk (E)  of predator awakening. The relative value of approach (vs. avoid) decreases with loss amount and threat level, and moreso under pessimistic assumptions. (Parameters: $\gamma = 1.0$)
\end{figure}

@@

As shown in Fig.~2, our model easily accommodates this result. Whereas optimistic agents pump until the marginal gain of a pump no longer offsets the chance of the balloon bursting, optimal choice under increasingly pessimistic (i.e., anxious) assumptions cashes out progressively earlier -- similar to empirical findings\citep{Maner2007, ramirez2015}. This is because it anticipates and avoids future errors in choice, which would otherwise result in the balloon popping. Our model can analogously explain other manifestations of biased approach-avoid conflict in anxiety, such as in the predator avoidance task\cite{fung2019}.

@@

Relatedly, the model can also capture findings of increased behavioral inhibition, measured as prolonged response times, in anxious individuals under threat\cite{bach2015}. In the behavioral inhibition task, participants seek tokens adjacent to a virtual "sleeping predator", which are all lost if the predator awakes. Though in this variant the risk of predation is constant throughout a trial (rather than increasing, as in the BART) the potential loss from capture still increases with each token collected. Bach \cite{bach2015} finds that participants are slower to collect tokens as this potential loss increases, and that this slowing is enhanced by subclinical anxiety. We can capture this effect in our model by noting that the relative value of approach compared to avoidance is reduced as potential loss and the risk that the predator awakes increase (Fig.~2d/e). These effects are amplified under more pessimistic (i.e., in our model, more anxious) assumptions about future actions. Thus, as before, anxious pessimism in our model produces greater and earlier choice of avoidance, analogous to earlier cash-out in the BART. To extend this effect to reaction times, we adopt the further, standard assumption that actions (here, approach) are slower when their relative value compared to alternatives is lower \citep{oud2016}. In this case, the model captures behavioral inhibition (slower responses as threat increases), and its enhancement by anxiety, as measured by Bach. (Note that we need not assume the coupling of reaction times to action value spread is due to difficulty in decision formation per se, which Bach argues against: it may, for instance, reflect Pavlovian initiation biases \cite{niv2007}.)

@@

Another laboratory phenomenon associated with anxiety is ``aversive pruning'' in planning \cite{Huys2012, Lally2017}. This refers to the idea that when evaluating future action trajectories in a sequential task like chess, people are resource-limited, cannot evaluate all possible options, and must selectively consider certain paths and neglect others. One proposal for how people accomplish this is aversive pruning \citep{Huys2012}, wherein choice sequences involving large losses are discarded from further evaluation. An example of aversive pruning is shown in Fig.~3a. Although the optimal choice in the decision tree from is to weather an initial large loss (e.g., -70) in order to reap the large gain that follows, people tend to disfavor this path suggesting they prune it and consequently neglect the later gain. The degree of such pruning correlates, depending on the study, with subclinical depressive \cite{Huys2012} or anxiety \cite{Lally2017} symptoms.

Our model predicts this result (Fig.~3b) as specifically linked to our model of anxious pessimism, though for a somewhat different reason than in Huys' original modeling. In our model, pessimistic (anxious) agents neglect large gains deeper in the tree, not because they fail to consider them (here we assume full evaluation of the Bellman equation), but because with increasing anxiety they increasingly expect the potential of choosing incorrectly afterward, thus failing to recoup the loss (and, mathematically, probabilistically pruning the better branches). Future research could use slight variants in the decision trees to tease apart these different interpretations, for example, by comparing decision trees that differ only in what follows the large initial loss. Under a model of aversive pruning, such a change should not impact the proportion of agents selecting the left branch; in contrast, our model predicts choice should parametrically increase with the extent of the amelioration.

@@

Of note in this respect, anxiety and depression are highly comorbid, with almost half of individuals with a lifetime depression diagnosis also diagnosed with an anxiety disorder \citep{kessler2015}. One notable proposal is that this association often (though by no means exclusively) arises longitudinally: in particular, that clinical anxiety precedes certain types of depression \citep{alloy1990, jacobson2014}. The idea, in brief, is that uncertainty in one's ability in the face of future threat results in anxiety and avoidance behaviors. Persistent avoidance, in turn, begets foregone reward, and ultimately to a belief that reward is unobtainable and subsequently depression. This informal story can be captured by simulations of learning in our model (Fig.~4) in environments like that of Fig.~1. Over the course of learning, the penumbra of negative value under pessimistic assumptions spreads gradually throughout the environment. This can in turn lead the agent to expect no reward and, also echoing the anergic symptoms of depression, forego action altogether. This last point in particular dovetails nicely with theoretical accounts of the anergic aspects of depression \citep{huys2015}, which point out that low experienced reward rates should in decision theoretic accounts lead to reduced response vigor \citep{niv2007} leading to a potentially self-reinforcing downward spiral.

In addition to suggesting one explanation for the comorbidity of anxiety and depression, our model also hints at a reason for the longevity and recurrence of anxiety disorders even with treatment. Because pessimistic expectations allow for threat value to spread to states and actions far antecedent of the primary danger (e.g., Fig.~1d), it would accordingly also take a great many steps of iterative learning to correct all these exaggerated appraisals of threat. Frustratingly, these biased estimates of value may still remain even after a misbelief in the efficacy of future action is corrected for in a course of therapy. This phenomenon (similar to failures of model-free RL algorithms to adjust to reward revaluation without extensive relearning; \cite{daw2005}) may offer at least a partial answer to a classic puzzle in pathological avoidance, i.e. why it is so resistant to extinction \cite{moutoussis2018}, and to the unfortunately high rates of anxiety recurrence following treatment\citep{pittig2018}.

@@

We are not the first to propose a formal theory of control in psychiatry using MDPs. Huys \& Dayan \cite{HuysDayan2009} also provided a computational account of learned helplessness through simple models of one-step action-outcome contingencies. Our accounts differ particularly in our exclusive focus on control in the sequential setting, which Huys \& Dayan did not address. Indeed, we propose that ultimately key to anxiety is precisely the way in which evaluation in sequential tasks is necessarily reliant on expectations about future choice and events. Similarly, research and modeling by Bishop and colleagues \citep{browning2015, gagne2018} also taking a decision theoretic approach has stressed the importance of uncertainty as a core feature of anxiety. Specifically, they have described uncertainty as inherently aversive in anxiety, and have presented models of how uncertainty may be increased in anxiety (e.g. aberrant processing of environmental volatility). The present work is compatible with deficits in processing uncertainty, and might instead be viewed as an attempt to unpack why uncertainty is aversive: because, in our view, it is resolved (i.e. marginalized) under pessimistic distributional assumptions. As for control, we extend this view to focus on how uncertainty is resolved in the sequential setting, and also to zero in on particular instances of uncertainty (about future actions, and some other options discussed next) and misbeliefs about them that, we argue, are particularly consequential.

@@

@@

The possibility of multiple anxious phenotypes, each characterized by unique but not mutually exclusive beliefs, suggests the need for behavioral assays designed to isolate and interrogate such biases. One such task is the free choice premium paradigm described above, which captures pessimism (or optimism) about one's own choices. An analogous task might measure pessimistic expectations about environmental state transition probabilities. For instance, this could be accomplished in a variant of sequential decision making tasks that require subjects to learn the transition structure of a multi-step decision tree \citep{glascher2010, lee2014} and make choices to gather rewards in it. Pessimistic expectations about environmental state transitions would bias choices in this type of task. Individually, these tasks could test our hypothesis that either sort of misbelief is associated with symptoms of anxiety; they could also be compared to one another (and to more detailed self-report assessments of beliefs about control or self-efficacy) to investigate potential heterogeneity across patients in the antecedent of anxiety.

@@

It is important to note that the present model may not describe all anxiety disorders with equal accuracy. Indeed, our analysis of pessimistic sequential evaluation is, by definition, a model of prospective cognition. Thus, the present results are more likely to accurately describe the anxiety disorders which primarily involve aberrations in future-oriented cognitive processes, such as generalized and social anxiety disorders. Naturally, the present model can account neither for the compulsive behaviors of obsessive-compulsive disorder (OCD) nor the memory disturbances of post-traumatic stress disorder (PTSD). That said, recent clinical studies suggest that diminished perceived control is a vulnerability factor common to all anxiety disorders, including OCD and PTSD \citep{gallagher2014a, gallagher2014b}. Importantly, aspects of other psychiatric disorders that involve future-oriented misbeliefs, worry, and avoidance behaviors (e.g. eating disorders\citep{konstantellou2011}) may similarly be well-described by the current account. Indeed, and much more speculatively, the model may also have implications for bipolar disorders. The onset of manic symptoms is associated with overoptimistic perceived control \cite{alloy2006}, and our same model and basic reasoning (though now envisioning excessively optimistic rather than excessively pessimistic beliefs) may help to explain how this bias may translate into risk-seeking behavior and dysreguated goal pursuit. As such, the present results are transdiagnostic and not limited to one particular diagnosis.

Finally, the relationship between evaluation, planning, and neural replay discussed above suggests potential future work that might help to bring this model into contact with the more memory-related aspects of PTSD. For instance, the same replay processes that can be used to evaluate actions can also, in theory, update predictive representations, cognitive maps, or models of the environment \citep{russek_predictive_2017}, such as the successor representation \cite{dayan1993, momennejad_successor_2017}. If so, similar biases in replay could result in not just aberrant avoidance behavior, but also progressive, aberrant remodeling of world models or cognitive maps, an observation which may connect to the rich and complex set of issues on memory involvement in PTSD.

@@

\textbf{Aversive Pruning} In the aversive pruning task \citep{Huys2012, Lally2017}, participants learn to navigate a six-state graphworld, where each state is directly connected to only two other states. Each state is associated with some reward (cost), and participants must plan trajectories through the state-space so as to maximize reward (minimize cost).
